Demystifying the Computer Science \\ PhD Admission in the US\\{\large A Handbook for International \emph{and Domestic} Students

### 2024-04-20 demystify.tex
updates

The focus of the admissions committee is to evaluate your background and interest in research to see if you \textbf{fit} into our program and can handle a PhD in CS, which is a \emph{research} degree!  To evaluate your profile, we consider the following key indicators.

@@

\begin{commentbox}[Vu:]
  If your grading system is not US standard or you are from a good school unknown outside your country, you can ask your reference writers to explain that in their letters.  For example, ``Bach Khoa'' are the top universities in Vietnam for STEM studies but few people outside Vietnam know about them.  So if you are from there, you should ask your reference writers to mention that.
\end{commentbox}

\paragraph{How to help your LOR writer} Your LoR writer can ask you to provide additional information to help them write your letter.  For example, they might ask about GPA, research and work experiences, papers (if any), or anything you want them to mention. You can also provide them with a draft of your SOP so that they can see what you are saying about yourself and can complement that with their own perspective.  It might also be a good idea to provide them with such information even when they do not ask, especially if you have not interacted with them much or have not done much research with them.

@@

There are various guides on writing SOP , and \href{https://cs-sop.org/}{many example statements} are available. In short, focus on your research motivation and convince us that you can achieve it through your experience, e.g., published papers. You should also back up your claims with concrete evidence. For example, if you say you have experience with teaching, then show what you did, e.g., undergrad TA or mentoring someone.  If you say you work on project X, then show some results (paper submitted (or even rejected), achieved X\% performance improvement over ..). Also, do not use more than 2 pages (you can have references on the third page).

@@

This is something easy to do, but is neglected by many
applicants: \textbf{customize the statement} for the school you're applying to,
e.g., why do you apply \emph{here}? provide names of professors who you're interested in (if they are not already in the adcom, your application might get forwarded to them for evaluation; and they might be interested in interviewing and recruiting you).
This shows that you're serious and have done homework on places you're applying to.
Adcom will look for this part (\autoref{sec:why-rejected}), so do not skip it.

>>>>

The focus of adcom is to evaluate your background and interest in research to see if you \textbf{fit} into our program and can handle a PhD in CS, which is a \emph{research} degree!  To evaluate your profile, we consider the following key indicators.

@@

\begin{commentbox}[Vu:]
  If your grading system is not US standard, or you are from a good school unknown outside your country, you can ask your reference writers to explain that in their letters.  For example, ``Bach Khoa'' are the top universities in Vietnam for STEM studies but few people outside Vietnam know about them.  So if you are from there, you should ask your reference writers to mention that.
\end{commentbox}

\paragraph{How to help your LOR writer} Your writer can ask you to provide additional information to help them write your letter.  For example, they might ask about GPA, research and work experiences, papers (if any), or anything you want them to mention. You can also provide them with a draft of your SOP so that they can see what you are saying about yourself and can complement that with their own perspective.  It might also be a good idea to provide them with such information even when they do not ask, especially if you have not interacted with them much or have not done much research with them.

@@

In your SOP, focus on research motivation and convince us that you can achieve it through your experience, e.g., published papers. You should also back up your claims with concrete evidence. For example, if you say you have experience with teaching, then show what you did, e.g., undergrad TA or mentoring someone.  If you say you work on project X, then show some results (paper submitted (or even rejected), achieved X\% performance improvement over ..). Also, do not use more than two pages (references can be on the third page). \href{https://cs-sop.org/}{Many SOP samples} are available.

@@

This is a simple task often overlooked by many applicants: \textbf{tailor your SOP} to the institution you're applying for,
e.g., why do you apply \emph{here}? Provide names of professors who you're interested in (if they are not already in the adcom, your application might get forwarded to them for evaluation; and they might be interested in interviewing and recruiting you).
This shows that you're serious and have done homework on places you're applying to.
Adcom will look for this part (\autoref{sec:why-rejected}).
